---
title: "p8105_hw5_KJW2175"
author: "Kennedy Wade"
date: "2025-11-05"
output: github_document
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(purrr)
library(p8105.datasets)

```


# Problem 1

```{r}
birthdays = sample(1:365, 5, replace = TRUE)

repeated_bday = length(unique(birthdays)) < 5

repeated_bday

bday_sim = function (n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)

repeated_bday = length(unique(birthdays)) < n_room

repeated_bday
  
}

bday_sim(20)

bday_sim_results = 
  expand_grid(
    bdays = 5:50,
    iter = 1:12500
  ) |>
  mutate(
    results = map_lgl(bdays, bday_sim)
  ) |>
  group_by(bdays) |>
  summarize(
    prob_repeat = mean(results)
  )
 
bday_sim_results |>
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line() +
labs(
    x = "bdays",
    y = "Probability of Shared Birthday",
    title = "Birthday Problem Simulation"
  )
```

Here, I drew random birthdays for a fixed group size and checked if any birthdays were duplicated. This code chunk returned a logical value based on that check. My output of FALSE means that in this random draw, no one shared a birthday. 

This plot shows the probability that at least two people in a group share the same birthday as the group size increases. When the group is small (fewer than 10 people), the probability of a shared birthday is close to zero. As the number of people increases, the probability that two people share a birthday rises. 


# Problem 2

## Defining a function to simulate one dataset and compute the t-test, using a for loop to repeat it 5000 times, and storing results in a list and combining them with bind_rows().
```{r}
sim_ttest = function(n = 30, mu = 0, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma)
  )
  
  test_res = t.test(sim_data$x, mu = 0)
  
  tibble(
    mu_hat = mean(sim_data$x),
    p_value = test_res$p.value
  )
}

output = vector("list", 5000)

for (i in 1:5000) {
  output[[i]] = sim_ttest()
}

sim_results = bind_rows(output)

```

## Looping through each mean value (0–6), running 5000 simulations for each mean, calculating the proportion of p-values < 0.05, and plotting power vs true mean to see assocation
```{r}
mu_values = 0:6      
n_sim = 5000        


sim_grid = expand_grid(
  mu = mu_values,
  iter = 1:n_sim
)

sim_results = sim_grid |> 
  mutate(
    results = map(mu, ~sim_ttest(n = 30, mu = .x, sigma = 5))
  ) |> 
  unnest(results)


power_results = sim_results |> 
  group_by(mu) |> 
  summarize(
    power = mean(p_value < 0.05),
    .groups = "drop"
  )


power_results |> 
  ggplot(aes(x = mu, y = power)) +
  geom_point() +
  geom_line() +
  labs(
    x = "True μ",
    y = "Power (Proportion of Rejected Nulls)",
    title = "Power of One-Sample t-Test vs True Effect Size"
  )
```


## Making a plot showing the average estimate of 𝜇̂ on the y axis and the true value of 𝜇 on the x axis and  overlaying the 2 averages only in samples for which the null was rejected on the y axis and the true value of 𝜇 on the x axis.
```{r}
true_means = 0:6
n_sim = 5000

sim_grid = expand_grid(mu = true_means, iter = 1:n_sim)

sim_results = sim_grid |>
  mutate(res = map(mu, ~ sim_ttest(n = 30, mu = .x, sigma = 5))) |>
  unnest(res) |>
  group_by(mu) |>
  summarize(
    avg_all = mean(mu_hat),
    avg_rejected = mean(mu_hat[p_value < 0.05]),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(avg_all, avg_rejected),
    names_to = "type",
    values_to = "avg_mu_hat"
  )

ggplot(sim_results, aes(x = mu, y = avg_mu_hat, color = type)) +
  geom_point() +
  geom_line() +
  labs(
    x = "True Mu",
    y = "Average Estimated Mu_hat",
    title = "Average Estimated Mu vs True Mu",
    color = "Simulation Type"
  )
```


# Problem 3

## describing the raw dataset 
```{r}
homicides = read_csv("homicide-data.csv")
```

The raw data contains information regarding individual case records with detailed information about each victim and incident. The dataset includes variables for unique case IDs, reporting dates, victim demographics (name, race, age, sex), geographic location (city, state, latitude/longitude coordinates), and case disposition status. Key variables for analysis are the disposition categories: "Closed by arrest" indicates solved cases, while "Closed without arrest" and "Open/No arrest" represent unsolved homicides.

## Creating a city_state variable, summarizing within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).
```{r}
city_homicides <- homicides |>
  mutate(city_state = str_c(city, ", ", state))

city_summary <- homicides |>
  mutate(city_state = str_c(city, ", ", state)) |>
  group_by(city_state) |>
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest")),
    .groups = "drop"
  )
```


## Baltimore, MD proportion test
```{r}
city_homicides |>
  filter(city_state == "Baltimore, MD") |>
  mutate(
    unsolved = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 1, 0)
  ) |>
  summarize(
    unsolved_count = sum(unsolved, na.rm = TRUE),
    total_count = n()
  ) |>
  reframe(
    broom::tidy(prop.test(unsolved_count, total_count))
  ) |>
  select(estimate, conf.low, conf.high)

```
 
 
## Running `prop.test` for each city 
```{r}
get_city_prop_tests <- function(data) {
  data |>
    mutate(
      unsolved = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 1, 0)
    ) |>
    group_by(city_state) |>
    summarize(
      unsolved_count = sum(unsolved, na.rm = TRUE),
      total_count = n(),
      .groups = "drop"
    ) |>
    mutate(
      test = map2(unsolved_count, total_count, ~ prop.test(.x, .y)),
      tidy = map(test, broom::tidy)
    ) |>
    unnest(tidy) |>
    select(city_state, estimate, conf.low, conf.high)
}

city_prop_tests <- get_city_prop_tests(city_homicides)
print(city_prop_tests)
```

## Ploting proportions with 95% CIs
```{r}
city_prop_tests |>
  mutate(city_state = reorder(city_state, estimate)
  ) |>
  ggplot(aes(x = city_state, y = estimate)) +
  geom_col(fill = "orange") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  coord_flip() +
  labs(
    x = "City",
    y = "Prop. Unsolved Homocides",
    title = "Proportion of Unsolved Homicides by City (95% CI)",
  ) +
  theme_minimal(base_size = 14)
```
